---
title: "child2"
format: docx
editor: visual
---

# Modelling

A linear model was fitted with the following idea: - A circadian cycle was fitted with a small degree-of-freedom natural spline - Time since the last event was modelled with an additional spline on top of the circadian rhythm

The results are far from perfect. Actual problems:

-   Only interaction terms were included in order to better interpret the effect of various foods (the blood sugar level always originating from the same value). This is particularly bad because there may be lingering effects from the previous food effects.
-   There are meals with few observations (because I ate some more after a short while). It is not appropriate to fit a large degree-of-freedom spline for them and 'sparser' splines should be specified (by hand in the model matrix).
-   A better way of looking at things would be to model both time since last meal and any documented activities in addition to one another (at this current setup, an activity 'stops' the modelling of the food effect).

While there are some definite issues with this setup, and the distribution of the residuals are not too good, but the resulting model explains \~75% of the total variability and produces plausable graphs for the food/circadian effects.

```{r}

# Outliers based on Cook distance
data_w_notes <- data_w_notes[c(-3604,-4623,-4624,-4626),]

mod_notes_matrix <- model.matrix( ~                   
                  notes * ns( note_time_scaled, 
                     Boundary.knots = c(0,1),
                     knots = seq(0,1,length.out=7)[2:6])
                + notes_activity * ns( note_time_scaled_activity, 
                    Boundary.knots = c(0,1),
                    knots = seq(0,1,length.out=4)[2:3])
                + ns( tod, df = 4)
                - 1
                , data_w_notes) %>%
  as.data.frame() %>%
  bind_cols(data_w_notes %>% 
              dplyr::select(value3), .)

new_colnames <- colnames(mod_notes_matrix)

# Handle special cases for first few columns
new_colnames[1:4] <- c("value3", "N_activity", "N_food", "N_none")

# Apply renaming for the rest of the columns
new_colnames[5:length(new_colnames)] <- 
  gsub("notes", "", new_colnames[5:length(new_colnames)])

# new_colnames[5:length(new_colnames)] <- 
#   gsub("\\(.*\\)", "", new_colnames[5:length(new_colnames)])


# Apply the new column names to the matrix
colnames(mod_notes_matrix) <- new_colnames

mod_notes_nice_colnames <- lm(value3 ~ . 
                , as.data.frame(mod_notes_matrix))

```


```{r}

df <- 6

mod_notes <- lm(value3 ~ 
                  notes * ns( note_time_scaled, 
                     Boundary.knots = c(0,1),
                     knots = seq(0,1,length.out=df+2)[2:(df+1)])
                # + notes_activity * ns( note_time_scaled_activity, 
                #     Boundary.knots = c(0,1),
                #     knots = seq(0,1,length.out=4)[2:3])
                + ns( time_note_max_activity,
                      Boundary.knots = c(1,7200),
                      knots = c(2000,4000))
                + ns( tod, df = 5)
                + fourier_var
                #+ fourier20[,1]
                , data_w_notes)

data_w_notes$resid <- mod_notes$residuals
data_w_notes$pr    <- predict(mod_notes)
data_w_notes$pr_std <- predict(mod_notes, 
                               newdata = data_w_notes %>% 
                                 mutate(tod = 1,
                                        notes_activity = 'none',
                                        note_time_scaled_activity = 0,
                                        fourier_var = 0))
data_w_notes$pr_stdf <- predict(mod_notes, 
                               newdata = data_w_notes %>% 
                                 mutate(
                                        notes_activity = 'none',
                                        note_time_scaled_activity = 0,
                                        fourier_var = 0))


summary(mod_notes)$r.squared
summary(mod_notes)$adj.r.squared

```

```{r}

data_w_notes %>%
  ggplot(aes(x = time_note/3600, y = resid)) +
    theme_bw() +
    geom_point(alpha = .1) +
    geom_smooth(method = "loess", span = .114)

```


```{r}


anova(mod_notes) %>% as.data.frame() %>% kableExtra::kbl()


```

```{r}

cooks.distance(mod_notes) %>% 
  sort(decreasing = TRUE) %>% 
  head(n = 10) %>% 
  as.data.frame() %>%
  `colnames<-`("Most extreme Cook's distances") %>%
  kableExtra::kbl()


plot(acf(mod_notes$residuals,lag = 100))
plot(pacf(mod_notes$residuals,lag = 100))

plot(mod_notes,1)
plot(mod_notes,2)
plot(mod_notes,3)

```

```{r}

data_w_notes %>%
  ggplot(aes(x=value3,y=pr)) +
    theme_bw() +
    geom_abline(slope=1,intercept=0,color = "salmon") +
    geom_point(size = .25, alpha = .2)

```

```{r}



```

```{r}

data_w_notes %>%
  filter( notes == "none",
          notes_full == 'none',
          time_note_activity == 0
    #time_note == 0
    ) %>%
  ggplot(aes(x = hms::hms(tod), y = pr_stdf)) +
    theme_bw() +
    geom_point() +
    geom_line() +
    labs( x = "Time of day",
          y = "Predicted value")


fasting_gluc_at_8h <- 
  data_w_notes %>%
    filter( notes == "none",
            notes_full == 'none',
            time_note_activity == 0,
            tod > hms::hms(seconds = 0, minutes = 50, hours = 7),
            tod < hms::hms(seconds = 0, minutes = 10, hours = 8),
      #time_note == 0
      ) %>%
    .$pr %>%
    mean

```

```{r}

pr <- expand.grid( notes = unique(data_w_notes$notes),
                   time_note = seq(0,43200,length.out = 1),
                   notes_activity = "none",
                   note_time_scaled_activity = 0,
                   time_note_max_activity = seq(1,3600*2,length.out=100),
                   tod = seq(1,3600*24,length.out = 1),
                   fourier_var = 0)

pr <- 
  left_join( pr, 
             notes_max_time %>% 
               group_by(notes) %>%
               arrange(desc(time_note)) %>%
               slice(1) %>%
               rename( time_note_max = "time_note"), 
             by= c("notes")) %>%
  mutate(note_time_scaled = time_note / time_note_max) %>%
  filter(time_note <= time_note_max ) %>%
  mutate(bin = cut(time_note_max, 
                   breaks = 6, 
                   #labels = c("Bin1", "Bin2", "Bin3"),
                   include.lowest = TRUE)) %>% 
  filter(notes == "none") %>% 
  mutate(notes=forcats::fct_drop(notes)) #%>%
  # left_join( y = data_w_notes %>% select(notes,notes_category) %>%
  #              group_by(notes) %>% slice(1),
  #            by = "notes")

pr$pr. <- predict(mod_notes, newdata = pr)

pr <- pr %>%
  filter( pr. > 0,
          pr. < 15)


pr %>%
  ggplot( aes( x = time_note_max_activity/3600, y = pr.
                   #,color = as.numeric(i)
                   #, group = notes
                   )) +
        theme_bw() +
        theme(legend.position = "none") +
        geom_line(color= "black", linewidth = 1.1) +
        scale_y_continuous(limits = c(2,12),
                           breaks = c(3.5,5,8,11)) +
        scale_x_continuous(
          limits = c(0,
                     ceiling(max(pr$time_note_max_activity)/3600 /3)*3),
          breaks = c(0,3,6,9,12)) +
        geom_hline(yintercept = pr$pr.[pr$time_note_max_activity == 0], 
                   color = "salmon4") +
        labs(x = "Time since activity(!) (hours)",
             y = "Predicted values",
             title = "")

```

```{r}

pr <- expand.grid( notes = unique(data_w_notes$notes),
                   time_note = seq(0,43200,length.out = 1),
                   notes_activity = "none",
                   note_time_scaled_activity = 0,
                   time_note_max_activity = seq(1,3600*2,length.out=1),
                   tod = seq(1,3600*24,length.out = 1),
                   fourier_var = seq(-1,1,length.out=100))

pr <- 
  left_join( pr, 
             notes_max_time %>% 
               group_by(notes) %>%
               arrange(desc(time_note)) %>%
               slice(1) %>%
               rename( time_note_max = "time_note"), 
             by= c("notes")) %>%
  mutate(note_time_scaled = time_note / time_note_max) %>%
  filter(time_note <= time_note_max ) %>%
  mutate(bin = cut(time_note_max, 
                   breaks = 6, 
                   #labels = c("Bin1", "Bin2", "Bin3"),
                   include.lowest = TRUE)) %>% 
  filter(notes == "none") %>% 
  mutate(notes=forcats::fct_drop(notes)) #%>%
  # left_join( y = data_w_notes %>% select(notes,notes_category) %>%
  #              group_by(notes) %>% slice(1),
  #            by = "notes")

pr$pr. <- predict(mod_notes, newdata = pr)

pr <- pr %>%
  filter( pr. > 0,
          pr. < 15)


pr %>%
  ggplot( aes( x = fourier_var, y = pr.
                   #,color = as.numeric(i)
                   #, group = notes
                   )) +
        theme_bw() +
        theme(legend.position = "none") +
        geom_line(color= "black", linewidth = 1.1) +
        # scale_y_continuous(limits = c(2,12),
        #                    breaks = c(3.5,5,8,11)) +
        scale_x_continuous(
          limits = c(-1,1)) +
        geom_hline(yintercept = pr$pr.[pr$time_note_max_activity == 0], 
                   color = "salmon4") +
        labs(x = "Time since activity(!) (hours)",
             y = "Predicted values",
             title = "")

```


```{r}

pr <- expand.grid( notes = unique(data_w_notes$notes),
                   time_note = union(seq(0,43200,length.out = 120),28800),
                   notes_activity = "none",
                   note_time_scaled_activity = 0,
                   time_note_max_activity = 1,
                   tod = seq(1,3600*24,length.out = 10),
                   fourier_var = 0)

pr <- 
  left_join( pr, 
             notes_max_time %>% 
               group_by(notes) %>%
               arrange(desc(time_note)) %>%
               slice(1) %>%
               rename( time_note_max = "time_note"), 
             by= c("notes")) %>%
  mutate(note_time_scaled = time_note / time_note_max) %>%
  filter(time_note <= time_note_max ) %>%
  mutate(bin = cut(time_note_max, 
                   breaks = 6, 
                   #labels = c("Bin1", "Bin2", "Bin3"),
                   include.lowest = TRUE)) %>% 
  filter(notes != "none") %>% 
  mutate(notes=forcats::fct_drop(notes)) #%>%
  # left_join( y = data_w_notes %>% select(notes,notes_category) %>%
  #              group_by(notes) %>% slice(1),
  #            by = "notes")

pr$pr. <- predict(mod_notes, newdata = pr)

pr <- pr %>%
  filter( pr. > 0,
          pr. < 15)

```

```{r}

pr %>%
  ggplot(aes(x = pr.)) +
  theme_bw() +
  geom_histogram(aes(y = ..density..), alpha = .5, fill = "salmon4", 
                 color = "red", binwidth = .1) +
  scale_x_continuous(limits = c(0, 15)) +
  geom_histogram(data = data_w_notes, 
                 mapping = aes(x = value3, y = ..density..),
                 fill = "deepskyblue", 
                 color = "deepskyblue3", 
                 alpha = .5, binwidth = .1) +
  labs( x = "",
        title = "Histograms of observed (blue) / predicted (red) values")

```



```{r}

make_single_plot <- function(i = 1) {
  notes_act <- unique(sort(pr$notes))[i]
  pr_act <- 
    pr %>%
      filter(tod <= 28801,
             tod >= 28800) %>%
      filter( notes %in% notes_act) 
  color. <- colorRampPalette(c("blue","green","red"))(length(unique(sort(pr$notes))))[i]
  out <- 
    pr_act%>%
      ggplot( aes( x = time_note/3600, y = pr.
                   #,color = as.numeric(i)
                   #, group = notes
                   )) +
        theme_bw() +
        theme(legend.position = "none") +
        geom_line(color=color., linewidth = 1.1) +
        scale_y_continuous(limits = c(2,12),
                           breaks = c(3.5,5,8,11)) +
        scale_x_continuous(
          limits = c(0,
                     ceiling(max(pr_act$time_note)/3600 /3)*3),
          breaks = c(0,3,6,9,12)) +
        geom_point(data = data_w_notes %>% 
                     filter( notes %in% notes_act), 
                   mapping = aes(x = time_note/3600, 
                                 y = pr_std + resid),
                   alpha = .5,
                   size = .5,
                   colr = color.) +
        geom_hline(yintercept = pr_act$pr.[pr_act$time_note == 0], 
                   color = "deepskyblue") +
        geom_hline(yintercept = fasting_gluc_at_8h, 
               color = "salmon4") +
        labs(x = "Time since meal (hours)",
             y = "Predicted values",
             title = notes_act)
  return(out)
  
}


```

```{r}

make_faceted_plot <- function(from = 1, to = 9) {
  
  plots <- list()

  for ( i in from:(from+8)) {
    plots[[as.character(i)]] <- make_single_plot(i = i)
  }
  plots <- plots[order(names(plots))]
  
  do.call(ggpubr::ggarrange, c(plots, list( ncol = 3,nrow = 3)))
}

```


```{r}

p <- 1
make_faceted_plot(p)
p <- p + 9
make_faceted_plot(p)
p <- p + 9
make_faceted_plot(p)
p <- p + 9
make_faceted_plot(p)
p <- p + 9
make_faceted_plot(p)
p <- p + 9
make_faceted_plot(p)


```

```{r}

sjPlot::tab_model(mod_notes_nice_colnames)

```
