---
title: "child2"
format: docx
editor: visual
---

# Calibration

I have calibrated my sensors via fingertip measurements in a haphazard way. xDrip+ automatically advances the measurements by 10 minutes, due to the subcutaneous readings' lag (I've verified this, not shown here). I applied the calibration for the first sensor only, as the second one is supposedly calibrated by default and seemed pretty stable (after a considerable warm-up period).

```{r}

d_bg %>%
  filter( is.na(sensor) == FALSE,
          bad == 0) %>%
  ggplot(aes(x=bg,y=bg_cbgm)) +
    theme_bw() +
    geom_abline(slope=1,intercept =0, color="red") +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    geom_smooth( se = FALSE, color = "deepskyblue", 
                 method = "loess", span = 1.25) +
    facet_wrap(facets="sensor") +
    labs( title = "Calibration points per sensor",
          y = "CGM value",
          x = "Stick value")

```

 

Final calibration values shown below.

```{r}

data %>%
  ggplot(aes(x=value2,y=value3)) +
  theme_bw() +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  labs(x = "Calibrated value from app",
       y = "Calibrated value based on model")

```

The calibration points along with their place in a given day are also shown (because I could).

```{r}

plots <- list()
dayz  <- unique(as.Date(data$time))
for ( i in 1:length(dayz)) {
  tim_act <- as_datetime(dayz[i])

  tim_lim <- c(tim_act, tim_act + 3600 * 24)
  
  plots[[as.character(tim_lim[1])]] <-
    data %>%
      ggplot(aes(x=time,y=value2)) +
       theme_bw() +
       geom_line() +
       geom_point() +
       geom_point(data = d_bg, mapping = aes( y = bg), color = "red") +
       geom_point(data = d_bg, mapping = aes( y = bg_cbgm), color = "green") +
       scale_x_datetime(limits = tim_lim)

}
plots <- plots[order(names(plots))]

do.call(ggpubr::ggarrange, c(plots, list( ncol = 6,nrow = 3)))

```

 \

# Raw results

 \

I've plotted my readings below. There is an apparent difference between the sensors but this also could be a calibration issue.

```{r}

data %>%
  ggplot(aes(x=time,y=value3)) +
    theme_bw() +
    geom_point() +
    geom_line() +
    facet_wrap(facets="sensor",scales = "free_x") +
    labs(y = "Glucose (mmol/l)")

```

Plotting the results onto a 24 hour window gives the below picture. This of course is the result of my eating habits which are shall we say hard to quantify.

```{r}

data %>%
  ggplot(aes(x=time_of_day,y=value3)) +
    theme_bw() +
    geom_line(alpha=.2) +
    geom_point() +
    geom_smooth() +
    geom_smooth(method = "loess",span=.1,color="green") +
    geom_point(data = d_bg, mapping = aes( y = bg), color = "red") +
    labs(x = "Time of day",
         y = "CGM reading")
    
```

# Multivariate approach

Tried to visualize the readings in the following way:

-   meals were logged, so the time since last meal could be determined at each timepoint
-   elevation in blood glucose were listed 45; 60; 90 minutes after meals, and
-   k-means clustering was attempted in order to identify potential interesting patterns

The patterns alas were proven to be moderately interesting, with some types of meals influencing blood glucose minimally or massively but overall, in a uniform pattern.

```{r}

mod2 <- kmeans(dat_mvt2[,2:4],centers = 3)

dat_mvt2$clust <- mod2$cluster

# 3D scatter plot
scatterplot3d::scatterplot3d(dat_mvt2[,2], dat_mvt2[,3], dat_mvt2[,4], 
              color = dat_mvt2$clust, pch = 19)

```

## Recurring oscillations

Anticipating the problem of high autocorrelation, and taking into consideration the pulsatile nature of hormone release in general, I investigated the results with a periodogram (searching for oscillations as a function of frequency). Though noisy, a pattern of oscillations every 20 minutes may be interesting. However, inserting an oscillating sine/cosine wave at this frequency did not improve the models (see later).

```{r}

##############
# Periodogram stuff

periodogr <- 
  data.frame( time =
                seq(
                  min(data_w_notes$time),
                  max(data_w_notes$time),
                  by = 60
                ))
periodogr$val <- f(periodogr$time)
periodogr_ts <- ts(periodogr)

periodogram_b <-
  spec.pgram(periodogr_ts, 
             spans = c(7,9),
             taper = 0.005,
             #fast = FALSE
             #demean = FALSE,
             #detrend = FALSE
             plot = FALSE
  )

data.frame( freq = periodogram_b$freq
            ,spec = periodogram_b$spec[,1]
) %>%
  ggplot(aes( x = 1/freq, y = spec)) +
  theme_bw() +
  scale_x_log10(limits = c(2,1500),
                breaks = c(10,20,30,60,120,240,1000,1440)) +
  scale_y_log10() +
  geom_line() +
  geom_vline(xintercept = 17.4, color = "salmon")

```

